---
title: "tutorial-RCT"
output: html_document
date: "2023-06-17"
---

Suppose we want to assess the effectiveness of a new teaching method on student performance. We randomly assign students to either the treatment group, which receives the new teaching method, or the control group, which receives the traditional teaching method. We will then compare the mean scores between the two groups.

Step 1: Install and Load the Package
```{r}

library(RCT)
library(randomizr)

```

 
Step 2: Generate Random Assignment
```{r}
set.seed(123)  # Set a seed for reproducibility
n <- 100  # Total number of students
prob <- 0.5  # Probability of assignment to treatment group

# Random assignment using complete_ra()
assignment <- complete_ra(N = n, prob = prob);assignment

```

## Step 3: Simulate Outcome Data

For demonstration purposes, let's simulate the outcome data. Suppose the outcome variable is a continuous score. We'll assume that scores follow a normal distribution with a mean of 70 in the control group and a mean of 75 in the treatment group.

```{r}

# Simulate outcome data
control_scores <- rnorm(sum(assignment == 0), mean = 70, sd = 5)
treatment_scores <- rnorm(sum(assignment == 1), mean = 75, sd = 5)

# Combine scores for the two groups
scores <- c(control_scores, treatment_scores)

```

Step 4: Descriptive Statistics and Balance Check

```{r}

# Descriptive statistics
mean_control <- mean(scores[assignment == 0])
mean_treatment <- mean(scores[assignment == 1])

# Check balance
covariate <- NULL  # No covariate in this example
balance_table( data = scores, treatment  = assignment )

```

Step 5: Treatment Effect Analysis

```{r}
# T-test to compare means
treatment_effect <- t.test(scores[assignment == 1], scores[assignment == 0])

# Extract relevant information
effect_size <- treatment_effect$estimate
p_value <- treatment_effect$p.value
conf_interval <- treatment_effect$conf.int

```

Step 6: Interpretation
```{r}
# Interpretation
if (p_value < 0.05) {
  conclusion <- "The new teaching method has a significant effect on student performance."
} else {
  conclusion <- "There is no significant difference in student performance between the two teaching methods."
}

# Print results
cat("Treatment Effect:", effect_size, "\n")
cat("p-value:", p_value, "\n")
cat("Confidence Interval:", conf_interval, "\n")
cat("Conclusion:", conclusion, "\n")
```

This example demonstrates a basic RCT analysis using the `randomizr` package. However, keep in mind that the analysis methods and interpretation may vary depending on the specific research question and the characteristics of your data.